### 2017-03-14 thesis/intro/intro.tex
Draft thesis completed; now edit the whole thing - tmp commit
Edited Intro

\chapter{Introduction (draft)}
\label{ch:intro}

@@

Obesity has been one of the major global problems for more than a decade, associated with many noncommunicable diseases such as diabetes, cardiovascular diseases and certain types of cancers \citep{WHO2014}.
In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category \citep{WHO2000}.

The number of overweight and obese people, in both adults and children, have risen in every country of the world, and the trend continues as we speak.
Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society \citep{Lim2012}.

@@

Obesity is defined as an abnormal or excessive fat accumulation that may impair health of that individual \citep{Garrow1988}.
One common and widely used approach to categorise individuals is by measuring the \gls{bmi} of the individual.
\gls{bmi} is a measurement based on the weight-to-height ratio of an individual that is used by clinicians and epidemiologists to classify adults into underweight, normal weight, overweight or obese categories.
The unit of \gls{bmi} is defined as the individual's weight in kilograms per square of the height in meters (kg/m$^2$).

\gls{who} (2014) have used \gls{bmi} measurements to define overweight and obesity in adults as an individual with \gls{bmi} $\geq$ 25 kg/m$^2$ and $\geq$ 30 kg/m$^2$, respectively.
For children under the age of 5, overweight and obesity were defined as weight-for-height greater than 2 standard deviation and 3 standard deviation above \gls{who} Child Growth Standards median, respectively.
For children aged between 5 to 19 years, overweight and obesity were defined as BMI-for-age greater than 1 standard deviation and 2 standard deviation above \gls{who} Growth Reference Standards median, respectively.

Although the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the individual's metabolic status remains controversial, many studies have used \gls{bmi} due to the ease of data collection compared to other measurements, such as waist-to-height ratio \citep{Lee2008, Gelber2008}.
Consequently, most of the publicly available clinical data have height and weight data from which one can calculate the patients' \gls{bmi}, and therefore able to collect the patients' obesity status more conveniently than, for example, waist-to-height ratio.

@@

\noindent
This equated to approximately 1 in every 14 people in the world classed as obese in 2014, and therefore, 1 in 14 people around the world were also at severe health risks associated with obesity.
It was also noted in the report that women were more likely to be obese than men, and that the prevalence of overweight and obesity increased as the income level of the countries increased \citep{WHO2014}.

@@

% TODO: Add western-specific and NZ specific statistics on obesity
% TODO: Add male/female specific stats
% TODO: Add Maori/PI specific stats

% Until these terrible trends stop, and even after the trends stop, the world will need a plan to live along with this ongoing global epidemic of obesity and its associated diseases.

@@

Obesity is caused by continuous intake of food without expending all of the energy gained.
The key to the problem is the control of food intake and energy expenditure, and how they are regulated in your body.
The role of hypothalamus is central to the control of satiety and many mutations that cause obesity affects the satiety signals to and/or from hypothalamus.
The hypothalamus receives both nueronal and hormonal inputs from the body and propagates the signal to the downstream neurons that ultimately affect the food intake of the individual \citep{Bell2005, Spiegelman2001}.

@@

Insulin, perhaps the most well-known hormone for its role in glucose homeostasis, also has an important role in satiety and food intake regulation.
Briefly, when insulin binds to the insulin receptor, \gls{irs} protein family is recruited and phosphorylated by the insulin receptor's intrinsic tyrosine kinase activity \citep{Saltiel2002}.
The phosphorylated \gls{irs} protein then activates \gls{pi3k}, which in turn activates its downstream targets, causing a rapid signal cascade through the cell \citep{Saltiel2002}.

The evidence of insulin's involvement in the satiety control was not particularly clear even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor, and that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
This was perhaps due to the fact that leptin and insulin used two different signalling pathways in the cell; leptin used the \gls{jak}/\gls{stat} signalling pathway, whereas insulin used the \gls{pi3k} pathway for signalling \citep{Ghilardi1996}.
This was puzzling to the researchers as there seemed to be no converging point of the two hromones in these pathways.

However, leptin, and subsequently insulin, was shown to activate ATP-dep\-endent K$^+$ channel in glucose-responsive neurons in the hypothalamus \citep{Spanswick1997, Spanswick2000}.
The activation of K$^+$ channel provided evidence for the two pathways to act similarly on the cell, at least for an acute response, suggesting a convergence in the pathway response by the two hormones, and therefore their role in satiety control.

@@

There are many environmental factors associated with the global rise in obesity.
Even though these environmental factors are probably the most significant contributor to the current global obesity epidemic, there are some individual-level, genetic factors that contribute to the overall increase in the proportion of the population that are obese.

There is much evidence of obesity being caused by individual-level genetic factors that make some individuals more predisposed to obesity than others.
Generally speaking, there are two classes of genetically-driven obesity: monogenic obesity and polygenic obesity \citep{Moustafa2013}. \\

\noindent
Monogenic obesity is where a single mutation in one of the genes that is directly involved in the satiety controlling pathway causes severe obesity in that individual.
Since these mutations are usually directly related to the key constituents in the pathway mentioned earlier (\cref{sub:physiological_mechanism_of_obesity}), the resulting phenotype from the mutation causes severe early-onset obesity.
However, although these mutations cause a devastating effect on the individual carrying the mutation, these mutations occur very rarely in the population.
% With that said though, the identification of these mutations have shed light on the mechanisms in which obesity is developed and it is worth mentioning some of these mutations to highlight the importance of the components involved in the obesity pathway.

An example of a mutation that causes severe early-onset obesity is a mutation in the leptin gene.
\citet{Montague1997} found a single nucleotide deletion mutation at position 133 of the leptin gene in children with early-onset obesity.
They further showed that the mutation caused a frameshift mutation in the protein, resulting in an inactive form of leptin that was not able to be secreted \citep{Montague1997}.
As a result, children with this mutation was constantly hungry due to the lack of anorexogenic signals, and became obese at a very early stage of their lives.
\\

% In close relation to this mutation is the mutation in the leptin receptor gene.
% \citet{Clement1998} have discovered a single nucleotide substitution mutation that led to the leptin receptor transcript that skipped exon 16, leading to a form of the receptor with no transmembrane or intracellular domain.
% Individuals with this rare mutation showed extreme hyperphagia and symptoms of early-onset obesity, even though their circulating leptin levels were present abundantly \citep{Clement1998}.

% The next couple of mutations are associated with the synthesis and the processing of \gls{pomc}.
% \citet{Krude1998} found single nucleotide mutations in the exonic region of the \gls{pomc} gene that affected the transcription of the $\alpha$-\gls{msh} region (and some other regions) of \gls{pomc}.
% \citet{Jackson1997} identified a mutation in the \gls{pc1}, an enzyme responsible for the processing of many prohormones and neuropeptides, including \gls{pomc}.
% In either case, the mutation resulted in the lack of $\alpha$-\gls{msh} in the neurons for satiety signalling and caused severe early-onset obesity.

% As mentioned earlier, even though these mutations are very rare and affects only those individuals that have one of these mutations, it had a profound impact in the field of obesity research and contributed significantly to the knowledge of the mechanism of how obesity develops.

\noindent
Polygenic obesity is where a combination of variants in multiple genes or multiple regions of the genome, together with environmental factors, causes obesity in the individual \citep{Moustafa2013}.
Unlike in monogenic obesity where a single mutation in a gene causes severe obesity, the variants in some of the genes or genomic regions involved in polygenic obesity have little effect on the individual's obesity status by itself, but in combination with many other variants and together with environmental factors, it has a serious obesogenic effect on the individual.

% Known as the ``thrifty gene'' hypothesis, \citet{Neel1962} suggested that the genes that were important in obesity have been around in the population for many generations, as it would have had a selective advantage in an environment where the population experienced starvation more frequently than the current generation.
% However, as food became more accessible, a subset of the population who were particularly good at storing fat and survive the starvation period ``overreacted'' to this rapid change in the environment \citep{Bell2005, Spiegelman2001}.
% These observations in the rise in obesity in subsets of population cannot be explained simply by external factors alone, and there is a significant genetic component involved in this phenomenon \citep{Bell2005}.

There have been many studies that have investigated the association of genes and/or genomic regions with obesity.
Of those, many of the \glspl{gwas} published in 2007 that showed strong evidence of the \gls{fto} gene with \gls{bmi} and obesity were of greatest significance \citep{Dina2007,Frayling2007,Gerken2007,Scuteri2007}.
Until these studies came out, there has been no common genes or genomic regions that associated with obesity that was reproducible and reliable \citep{Frayling2007}.
Thus, these studies were the first evidence of common variants that contributed to the polygenic obesity phenotype in the population.

% Discovery of the \gls{fto} gene has lead to extensive research on the gene, and the mechanism of action by the \gls{fto} variant has been recently uncovered  from these efforts \citep{Claussnitzer2015a,Smemo2014a}.
% The association between obesity and genetic variations, and their mechanisms that lead to obesity, has not been fully elucidated due to its complexity, but it may not be long before they are fully uncovered.

@@

Even though some individual-level factors are involved with obesity, it is clear that these individual-level factors cannot be accounted for the global epidemic of obesity that the world currently observes.
For this reason, obesity is thought to be caused mainly by the environmental factors that act at a population-level.\\

\noindent
The first key factor that contirbutes to the global rise in the proportion of the population that is obese is the availability and abundance of certain food in the country.
With the help from government trade policies, food and other goods have become easier to trade between the country, helping the economic growth of many countries around the world \citep{Kearney2010}.
This led to greater abundance of food in general, greater number of choices of food and shifted the overall nutritional status of the country \citep{Malik2013}.
This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in population-wide rise in obesity.
% This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in a different, much larger problem -- obesity.

As an example, in the \gls{usa}, the cost of corn and soy is low due to the fact that they are the raw ingredients for most processed food and beverages, such as high fructose corn syrups used to sweeten many soft drinks in the \gls{usa} \citep{Malik2013}.
Moreover, corn and soy are also the main food for the livestock, which results in lower prices for meat \citep{Malik2013}.
On the other hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production \citep{Malik2013}.

As a result, more people are inclined to consume food that are cheap, have little nutritional value and high in energy, than the food that are expensive, nutritional and healthy alternatives.
Thus, the population is more likely to become obese by making poor food choice.\\

\noindent
The second factor to consider is the behavioural changes led by the urbanisation of the population.
% Urbanisation is defined as the gradual increase in the proportion of people living in urban areas.
As more people take up urban lifestyle, they are exposed to the new environment where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas.

Although there are many advantages involved with urbanisation, such as access to developed health care systems, education and advanced technologies, the lifestyle change also impose negative health behaviour that ultimately lead to positive energy balance and therefore obesity \citep{Malik2013}.
Lack of physical activity is one of the obvious components that is linked with the urbanisation of the population.
% Currently, the recommended amount of physical activity to keep an individual fit and healthy is $\geq$ 150 min of moderate-intensity aerobic physical activity per week, or $\geq$ 30 min of physical activity on most days of the week \citep{Pate1995, WHO2010}.
In many urban cities, the recommended level of physical exercise is difficult to achieve and maintain due to the shift towards the sedentary behaviours that are encouraged by the developed transport system, automated household chores, and indoor entertainment \citep{Malik2013}.
\\
% Urbanisation by itself cannot be accounted for the lack of exercise in the population, but it is true that urbanisation makes the bar higher for many people to take up the recommended level of fitness and maintain it. \\

\noindent
The last factor to be mentioned is the income and socioeconomic status of the population.
As the average income increases, more people are able to buy food and are likely to take up the sedentary lifestyle.

From this, one would expect the high-income group to have the largest impact from obesity, but on the contrary to this expectation, the biggest impact from obesity is seen in the low- and middle-income groups of the population \citep{Malik2013}.
Access to the developed health care system will allow weight maintenance of the population, but this is likely to be of most benefit for the high-income group in the population.
The reason for this, quite simply, is the cost associated with the visit to the health care system, as high-income group is more likely to make a consistent and periodic visit than the low or middle-income group of the population \citep{Malik2013}.

So, for the low- and middle-income groups, the rise in average income in these groups allow the people to have access to food and activities that promotes sedentary bahaviours, but the limited access to the health care system, education and recreational activities that promote physical activites ultimately lead to positive energy balance \citep{Malik2013}.
In contrast, the high-income group of the population have better access to the facilities that promote weight maintenance, thus have less of an impact compared with the low- and middle-income groups of the population.
\\

% \noindent
% The last factor, which ties in strongly with the first key factor, is the choice of diet made by the population.
% The amount of low-quality food have increased significantly in many rapidly growing countries, as these high-energy, low-nutritional food are cheaper and more profitible than the natural produce that have higher nutritional value \citep{Kearney2010}.
% As a result, more people tend to purchase these low-quality food more often, and consume more energy than they are required.

% However, many people eat responsibly and avoid these low-quality diet and prevent overconsumption of such food, but the overconsumption of these food is not a big problem as such, rather the overall quality of the food \citep{Malik2013}.
% These food are usually high in refined sugar and unhealthy fat, highly processed, contain very little nutritional value and are low in fibre.
% As a result, consumption of these food increases the overall energy intake and achieve no nutritional benefits from the food.

% Whether it is by choice or not, consumption of these low-quality diet leads to obesity and obesity-related diseases. \\

\noindent
All of these population-level environmental factors in synergy with the individual-level genetic factors have contributed to the global epidemic of obesity.

% \subsection{Impact of obesity}
% \label{sub:impact_of_obesity}

% As mentioned at the beginning of the current section, obesity has a devastating effect on the health of the population in a global scale.
% Overweight and obesity have been a risk factor for many diseases, including type 2 diabetes, cardiovascular diseases, hypertension, and many types of cancers \citep{Franks2010, Guh2009, Yach2006}.
% As \citet{Yach2006} mentions in their paper, ``overweight and obesity have become to diabetes what tobacco is to lung cancer'', and it is important to understand that obesity has significant risks and comorbidities associated with it.

% Although adulthood obesity is clinically important for risk assessment of diseases, childhood obesity has been focussed more heavily in recent years,  as it has greater public health implications.
% There are many immediate consequences that are related to childhood obesity, but the diseases that develop as these children grow up are more critical than the earlier effects.
% A paper by \citet{Must1999} reviewed many of the clinical studies done in the last several decades and summarised the risks associated with childhood obesity.
% Among the many risks associated with obesity, obese children, compared to lean children, had over 8.5-fold increased risk of developing hypertension as adults \citep{Must1999}.
% On top of this, the relative risks for all-cause mortality and coronary heart disease-specific mortality were 1.5 and 2.5, respectively \citep{Must1999}.

% In a more recent study by \citep{Franks2010}, they have shown that childhood obesity (as well as glucose intolerance and hypertension in childhood) was strongly associated with premature deaths from endogenous causes in American Indian population from Arizona (incidence rate ratio of 1.4 per unit of \gls{bmi} z-score).

% Health related consequences of obesity are clear from many epidemiological studies, but it is noteworthy that obesity also has an economical impact on the population as well.
% In the \gls{usa}, within the span of 5 years, the medical cost due to diabetes have risen from \$44 billion to \$92 billion \citep{Yach2006}.
% Although not all diabetes are directly caused by obesity, obesity would have assisted in the rise in the costs associated with diabetes.
% In addition to this,
% The costs that are associated with the treatments of comorbidities of obesity affects the household incomes, increased premature retirement and unemployment, which could have a significant economical impact on the society over time \citep{Yach2006}.

% As more of the younger generation is gaining weight and becoming obese, having the risks and disease burden associated with obesity at such early age are not ideal (both economically and epidemiologically) and poses a threat to the growing population that are becoming more obese.

% \subsection{Preventive measures of obesity}
% \label{sub:preventive_measures_of_obesity}

% It is apparent from previous sections that obesity has become a serious problem that the world needs to prioritise to solve immediately.
% Even though obesity has been regarded as one of the many global problems to solve, the progress has been slow.

% \citet{Sacks2009} proposed an \gls{opa} framework that was based on the framework developed by \citet{WHO2006}.
% In this framework, \citet{Sacks2009} have added extra features so that the policies were targeted to impact three levels of the underlying environment of obesity: the socio-ecological environment (upstream), lifestyle/behavioural changes (midstream), and health services support (downstream).

% The upstream approach targets the underlying economic, social and physical norms that the population is currently in.
% By targeting the broad socioeconomic environment that the population lives in, such as taxation, education, employment and housing \citep{Sacks2009}.
% In doing so, it is possible to have a significant change in the population behaviours and eventually improve the health outcomes of the population.

% The midstream approach targets the population or subpopulation behaviour change to improve the eating and physical activity behaviours in these groups.
% Unlike in the upstream approach, the midstream approach aims to change the behaviour by implementing social marketing and programmes that encourage such changes \citet{Sacks2009}.
% Again, the result of this approach will be an improvement in the overall positive behaviour of the population towards healthy eating and physical activity.

% The downstream approach targets to support the health services and clinical interventions.
% Unlike upstream and midstream approaches, downstream approach generally acts on an individual level, whereby the focus is on the management of the health of an individual and/or the individual's family \citet{Sacks2009}.
% Although not as efficient as upstream or midstream approaches, downstream approach will help improve the habits and behaviours to promote weight management.\\

% \noindent
% In the recent years, obesity has been declared as a global priority \citep{Malik2013}.
% Furthermore, comprehensive global monitoring framework, multi-sectoral actions and strengthening of the national policies have been implemented to better control the obesity epidemic \citep{Malik2013}.
% In many countries, the nutritional labelling on products have been adjusted to include the ingredients that have negative impact on obesity (for example, \textit{trans} fat), and even reduce or remove these ingredients from food \citep{Malik2013}.

% Food marketing can also be a problem in a highly competitive society that we live in.
% To sell their products that are high in fat, sugar and portion size, food industries market and advertise their products to the population, especially young children, in order to stay ahead in their market place \citep{Chan2010, Malik2013}.
% Even though there is a strong evidence of food marketing influencing the eating behaviours and food preferences in children, very few countries have been able to reduce these marketing and advertising strategies of unhealthy food \citep{Hawkes2007a, Malik2013}.

% In contrast to these upstream/midstream approaches, there are some clinical interventions and treatments that help manage weight in an individual.
% One of these therapies is bariatric surgery for obese patients, especially morbidly obese patients \citep{Buchwald2004}.
% Through their meta-analysis, \citet{Buchwald2004} showed that bariatric surgery was able to reduce many of the the comorbidities associated with obesity.
% The fact that these improvements and complete resolution of the comorbidities in high proportion of patients that underwent bariatric surgery shows that, even though it acts on an individual level, downstream approaches such as bariatric surgery may be a highly effective, but not an efficient, method of preventing obesity in the population.\\

% \noindent
% It is evident that the most efficient way to reduce the proportion of the population that are obese is to implement national policies that affect the upstream or midstream factors.
% These policies, although they are efficient at targeting the underlying problem, may take a long time to make a significant effect.
% On the otherhand, individual therapeutic plans that affect the downstream factors may be very effective in achieving the desired goal of weight loss, but it is very resource intensive and upscaling this approach would be practically impossible.

% The management of obesity at an early stage with midstream and/or downstream approaches while the upstream approach takes effect would be the key to success in controlling for the global obesity epidemic.

@@

% What is cancer.
\begin{quote}
	\textit{
	``Cancer is a generic term for a large group of diseases that can affect any part of the body \ldots{}
	One defining feature of cancer is the rapid creation of abnormal cells that grow beyond their usual boundaries, and which can then invade adjoining parts of the body and spread to other organs \ldots{}
	referred to as metastasizing.
	Metastases are the major cause of death from cancer.''
	\citep{WHO2016}
	}
\end{quote}

@@

% TODO: Add western-specific and NZ specific statistics on cancer
% TODO: Add male/female specific stats
% TODO: Add Maori/PI specific stats

Cancer is considered as one of the leading causes of deaths around the world, accounting for 8.2 million deaths in 2012 \citep{WHO2014}.
More than half of cancer deaths are caused by lung, breast, colorectal, stomach and liver cancers \citep{WHO2014}.
The leading cause of deaths in high-income countries is lung cancer in both men and women, then colorectal cancer and breast cancer for men and women, respectively \citep{WHO2014}.
Although the levels of mortality vary among different countries, cervical, liver and stomach cancers make up a large proportion of cancer deaths among low- and middle-income countries \citep{WHO2014}.
\\

\noindent
\citet{Bray2013} estimated the global prevalence of cancer in adult population in 2008.
Since there is no clear definition for cancer prevalence, \citet{Bray2013} used \gls{ldp} as an estimate for cancer prevalence, and measured the incidence of cancer during the period between 2004 and 2008 (5 years).
% measuring the incidence of cancer during the period between 2004 and 2008 (5 years) who were still alive at the end of 2008.
From these estimates, Eastern Asia had the greatest 5-year prevalence of cancer, with approximately 7 million people with cancer in 2008, followed by North America and Western Europe with approximately 4.7 and 3 million people with cancer \citep{Bray2013}.
When the population of these regions was taken into account and the proportion of cancer prevalence calculated, Western Europe had the highest prevalence proportion of 1.9\%, followed closely by Australia/New Zealand with 1.82\% and North America with 1.7\% \citep{Bray2013}.

Looking at the prevalence of cancer by specific types, breast cancer was the most prevalent type of cancer with 5.2 million cases in 2008 \citep{Bray2013}.
Colorectal and prostate cancers were second and third most prevalent cancer types, with approximately 3.2 million cases for  both of them \citep{Bray2013}.
\citet{Bray2013} noted that female-specific cancer types were the most prevalent types of cancer in 75\% of the countries in the world.
Furthermore, the prevalence of breast, cervix or prostate cancer is the highest in almost 95\% of the countries \citep{Bray2013}.
Although these estimates provide a relatively crude measure of cancer prevalence, \citet{Bray2013} provided a useful overview of what the prevalence of cancer is like around the world.\\

\noindent
The total number of deaths by non-communicable diseases (including cancer) is predicted to increase, reaching 52 million deaths by 2030 \citep{WHO2014}.
With the prevalence of cancer at such high levels around the world, and with cancer mortality increasing in the future, cancer is a serious burden that the world has to face.

>>>>

\chapter{Introduction}
\label{ch:intro}

@@

Obesity has been one of the major global problems for more than a decade, associated with many noncommunicable diseases such as diabetes, cardiovascular diseases and certain types of cancers \citep{WHO2014}.
In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category \citep{WHO2000}.
The number of overweight and obese people, in both adults and children, have risen in every country of the world and the trend will likely to continue in the future.
Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society \citep{Lim2012}.

@@

Obesity is defined as an abnormal or excessive fat accumulation that may impair health of that individual \citep{Garrow1988}.
One common and widely used approach to categorise individuals is by measuring the \gls{bmi} of the individual.
\gls{bmi} is a measurement based on the weight-to-height ratio of an individual that is often used by clinicians and epidemiologists to classify adults into underweight, normal weight, overweight or obese categories.
The unit of \gls{bmi} is defined as the weight in kilograms per square of the height in meters (kg/m$^2$).

\gls{who} (2014) have used \gls{bmi} measurements to define overweight and obesity in adults as individuals with \gls{bmi} $\geq$ 25 kg/m$^2$ and $\geq$ 30 kg/m$^2$, respectively.
For children under the age of 5, overweight and obesity were defined as weight-for-height greater than 2 standard deviation and 3 standard deviation above \gls{who} Child Growth Standards median, respectively.
For children aged between 5 to 19 years, overweight and obesity were defined as BMI-for-age greater than 1 standard deviation and 2 standard deviation above \gls{who} Growth Reference Standards median, respectively.

However, the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the individual's metabolic status remains controversial.
In fact, many studies have suggested the use of other measurements such as waist-to-height and waist-to-hip ratios to classify the patients into different categories, as these measurements were better predictors of the obesity related diseases than \gls{bmi} \citep{Dalton2003,Gelber2008,Lee2008}.
With that said, \gls{bmi} is still widely used due to the ease of data collection compared to other measurements and, as a result, most of the publicly available clinical data have  height and weight data of the patients from which one can calculate the patients' \gls{bmi}.

@@

\noindent
This equated to approximately 1 in every 14 people in the world classed as obese in 2014, and therefore, 1 in 14 people around the world were also at severe health risks associated with obesity.
The highest prevalence of overweight and obesity was in the American regions, where 61\% and 27\% of the population were overweight and obese, respectively \citep{WHO2014}.
It was also noted in the report that women were more likely to be obese than men, and that the prevalence of overweight and obesity increased as the income level of the countries increased \citep{WHO2014}.

@@

In \gls{nz}, obesity is a serious problem as approximately one third (31.6\%) of the adults in \gls{nz} were classified as obese in 2015/2016 \citep{Health2016}.
The Pacific and M\=aori populations were affected the most, where 67\% and 47\% of the adults, respectively, were obese \citep{Health2016}.
\gls{nz} not only have a high prevalence of obesity compared to many of the other countries of the world, but also have a unique health problem that affects the Pacific and M\=aori populations more than the rest of the \gls{nz} population.

% TODO: Add western-specific statistics on obesity (wang2007)
% TODO: Add male/female specific stats

@@

Obesity is caused by continuous intake of food without expending all of the energy gained from the food.
The key to the problem is the control of food intake and energy expenditure, and how they are regulated in the body.
The role of hypothalamus is central to the control of satiety and many of the mutations that cause obesity affects the satiety signals to and/or from hypothalamus.
The hypothalamus receives both nueronal and hormonal inputs from the body and propagates the signal to the downstream neurons that ultimately affect the food intake of the individual \citep{Bell2005, Spiegelman2001}.

@@

Insulin, perhaps the most well-known hormone for its role in glucose homeostasis, also has an important role in satiety and food intake regulation.
Briefly, when insulin binds to the insulin receptor, \gls{irs} protein family is recruited and phosphorylated by the  intrinsic tyrosine kinase activity of the insulin receptor \citep{Saltiel2002}.
The phosphorylated \gls{irs} protein then activates \gls{pi3k}, which in turn activates its downstream targets, causing a rapid signal cascade through the cell \citep{Saltiel2002}.

The evidence of the involvement of insulin in the satiety control was not particularly clear even though there were strong evidence of obesity with neuron-specific deletion of insulin receptor, and the fact that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
This was perhaps due to the fact that leptin and insulin used two different signalling pathways in the cell; leptin used the \gls{jak}/\gls{stat} signalling pathway, whereas insulin used the \gls{pi3k} pathway for signalling \citep{Ghilardi1996}.
However, leptin, and subsequently insulin, was shown to activate ATP-dep\-endent K$^+$ channel in glucose-responsive neurons in the hypothalamus \citep{Spanswick1997, Spanswick2000}.
The activation of K$^+$ channel provided evidence for the two pathways to act similarly on the cell, at least for an acute response, suggesting a convergence in the pathway response by the two hormones, and therefore their role in satiety control.

@@

Even though environmental factors are probably the most significant contributor to the current global obesity epidemic, there are some individual-level, genetic factors that contribute to the overall increase in the proportion of the population that are obese.
There is much evidence of obesity being caused by individual-level genetic factors that make some individuals more predisposed to obesity than others.
Generally speaking, there are two classes of genetically-driven obesity: monogenic obesity and polygenic obesity \citep{Moustafa2013}.

Monogenic obesity is where a single mutation in one of the genes that is directly involved in the satiety controlling pathway causes severe obesity in that individual.
Since these mutations are usually directly related to the key constituents in the pathway mentioned earlier (\cref{sub:physiological_mechanism_of_obesity}), the resulting phenotype from the mutation causes severe early-onset obesity.
However, although these mutations cause a devastating effect on the individual carrying the mutation, these mutations occur very rarely in the population.

An example of a mutation that causes severe early-onset obesity is a mutation in the leptin gene.
\citet{Montague1997} found a single nucleotide deletion mutation at position 133 of the leptin gene in children with early-onset obesity.
They further showed that the mutation caused a frameshift mutation in the protein, resulting in an inactive form of leptin that was not able to be secreted \citep{Montague1997}.
As a result, children with this mutation was constantly hungry due to the lack of anorexogenic signals, and became obese at a very early stage of their lives.
Though these mutations are very rare and affects only those individuals that have one of these mutations, it had a profound impact in the field of obesity research and contributed significantly to the knowledge of the mechanism of the development of obesity.

Polygenic obesity is where a combination of variants in multiple genes or multiple regions of the genome, together with the environmental factors, causes obesity in the individual \citep{Moustafa2013}.
Unlike in monogenic obesity where a single mutation in a gene causes severe obesity, the variants in some of the genes or genomic regions involved in polygenic obesity have little effect on the obesity status of the individual by itself, but in combination with many other variants and together with environmental factors, it has a serious obesogenic effect on the individual.

There have been many studies that have investigated the association of genes and/or genomic regions with obesity.
Of those, many of the \glspl{gwas} published in 2007 showed strong evidence of the \gls{fto} gene with \gls{bmi} and obesity  \citep{Dina2007,Frayling2007,Gerken2007,Scuteri2007}.
Until these studies came out, there has been no common genes or genomic regions that associated with obesity that was reproducible and reliable \citep{Frayling2007}.
Thus, these studies were the first evidence of common variants that contributed to the polygenic obesity phenotype in the population.

@@

It is clear that the individual-level factors cannot be accounted for the global epidemic of obesity that the world currently observes.
For this reason, obesity is thought to be caused mainly by the environmental factors that act on a population-level.
The first key factor that contributes to the global rise in the proportion of the population that is obese is the availability and abundance of certain food in the country.
With the help from the government trade policies, food and other goods have become easier to trade between countries, helping the economic growth of many countries around the world \citep{Kearney2010}.
This led to greater abundance of food in general, greater number of choices of food and shifted the overall nutritional status of the country \citep{Malik2013}.
This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in the population-wide rise in obesity.

As an example, in the \gls{usa}, the costs of corn and soy are low due to the fact that they are the raw ingredients for most processed food and beverages, such as high fructose corn syrups used to sweeten many soft drinks in the \gls{usa} \citep{Malik2013}.
Moreover, corn and soy are also the main food for the livestock, which results in lower prices for meat \citep{Malik2013}.
On the other hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production \citep{Malik2013}.

As a result, more people are inclined to consume food that are cheap, have little nutritional value and high in energy, than the food that are expensive, nutritional and healthy alternatives.
Thus, the population is more likely to become obese by making poor food choice.

The second factor to consider is the behavioural changes led by the urbanisation of the population.
As more people take up urban lifestyle, they are exposed to the new environment where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas.

Although there are many advantages involved with urbanisation, such as access to developed health care systems, education and advanced technologies, the lifestyle change also impose negative health behaviour that ultimately lead to positive energy balance and therefore obesity \citep{Malik2013}.
Lack of physical activity is one of the obvious components that is linked with the urbanisation of the population.
In many urban cities, the recommended level of physical exercise is difficult to achieve and maintain due to the shift towards the sedentary behaviours that are encouraged by the developed transport system, automated household chores, and indoor entertainment \citep{Malik2013}.

The last factor to be mentioned is the income and socioeconomic status of the population.
As the average income increases, more people are able to buy food and are likely to take up the sedentary lifestyle.
From this, one would expect the high-income group to have the largest impact from obesity, but on the contrary to this expectation, the biggest impact from obesity is seen in the low- and middle-income groups of the population \citep{Malik2013}.
Access to the developed health care system will allow weight maintenance of the population, but this is likely to be of most benefit for the high-income group in the population.
The reason for this, quite simply, is the cost associated with the visit to the health care system, as high-income group is more likely to make a consistent and periodic visit than the low or middle-income group of the population \citep{Malik2013}.

So, for the low- and middle-income groups, the rise in average income in these groups allow the people to have access to food and activities that promotes sedentary bahaviours, but the limited access to the health care system, education and recreational activities that promote physical activites ultimately lead to positive energy balance \citep{Malik2013}.
In contrast, the high-income group of the population have better access to the facilities that promote weight maintenance, thus have less of an impact compared with the low- and middle-income groups of the population.
All of these population-level environmental factors in synergy with the individual-level genetic factors have contributed to the global epidemic of obesity.

@@

\begin{quote}
	\textit{
	``Cancer is a generic term for a large group of diseases that can affect any part of the body \ldots{}
	One defining feature of cancer is the rapid creation of abnormal cells that grow beyond their usual boundaries, and which can then invade adjoining parts of the body and spread to other organs \ldots{}
	referred to as metastasizing.
	Metastases are the major cause of death from cancer.''
	\citep{WHO2016}
	}
\end{quote}

@@

Cancer is considered as one of the leading causes of deaths around the world, accounting for 8.2 million deaths in 2012 \citep{WHO2014}.
More than half of cancer deaths are caused by lung, breast, colorectal, stomach and liver cancers \citep{WHO2014}.
The leading cause of deaths in high-income countries is lung cancer in both men and women, then colorectal cancer and breast cancer for men and women, respectively \citep{WHO2014}.
Although the levels of mortality vary among different countries, cervical, liver and stomach cancers make up a large proportion of cancer deaths among low- and middle-income countries \citep{WHO2014}.

\citet{Bray2013} estimated the global prevalence of cancer in adult population in 2008.
Since there is no clear definition for cancer prevalence, \citet{Bray2013} used \gls{ldp} as an estimate for cancer prevalence, and measured the incidence of cancer during the period between 2004 and 2008 (5 years).
From these estimates, Eastern Asia had the greatest 5-year prevalence of cancer, with approximately 7 million people with cancer in 2008, followed by North America and Western Europe with approximately 4.7 and 3 million people with cancer \citep{Bray2013}.
When the population of these regions was taken into account and the proportion of cancer prevalence calculated, Western Europe had the highest prevalence proportion of 1.9\%, followed closely by Australia/New Zealand with 1.82\% and North America with 1.7\% \citep{Bray2013}.

Looking at the prevalence of cancer by specific types, breast cancer was the most prevalent type of cancer with 5.2 million cases in 2008 \citep{Bray2013}.
Colorectal and prostate cancers were second and third most prevalent cancer types, with approximately 3.2 million cases for  both of them \citep{Bray2013}.
\citet{Bray2013} noted that female-specific cancer types were the most prevalent types of cancer in 75\% of the countries in the world.
Furthermore, the prevalence of breast, cervix or prostate cancer is the highest in almost 95\% of the countries \citep{Bray2013}.
Although these estimates provide a relatively crude measure of cancer prevalence, \citet{Bray2013} provided a useful overview of the prevalence of cancer  around the world.
